---
title: "Practical 6 Answers"
subtitle: "GLMs: Logistic regression"
author: "Aysha Flueck"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: false
    link-external-newwindow: true
    link-external-icon: true
editor: source
---

```{r}
#| label: setup
#| message: false
#| warning: false

# Load the packages you need for this practical.
library(readr)
library(dplyr)
library(ggplot2)
library(tibble)
library(tidyr)
library(naniar)

titanic <- read.csv("data/titanic.csv")
```

## Titanic survival data

### Question 1
*Read the data in using read_csv and inspect for missingness and other data irregularities.*
```{r}
glimpse(titanic)

titanic |> filter(cabin == "") |> count()
titanic |> filter(boat == "") |> count()
titanic |> filter(home.dest == "") |> count()

```

`cabin`, `boat`, and `home.dest` all have missing entries whcih have been incorrectly coded as blank rather than `NA`.

```{r}
titanic <- titanic |>
  mutate(
    cabin = na_if(cabin, ""),
    boat = na_if(boat, ""),
    home.dest = na_if(home.dest, ""))
```


```{r}
miss_var_summary(titanic)
vis_miss(titanic)
gg_miss_upset(titanic)

titanic |> group_by(survived) |>
  miss_var_summary() |>
  filter(n_miss > 0)

titanic |> group_by(pclass) |>
  miss_var_summary() |>
  filter(n_miss > 0)
```

`body` and `boat` have a lot of missingness (1188 and 823), which is to be expected by construction. `boat` is almost exclusively recorded for survivors, and `body` only for the recovered non-survivors. 

- `cabin` is missing 1014, mostly for 2nd and 3rd class passengers.
- `home.dest` is missing 564.
- `age` is missing 263.

### Question 2
*Tidy the data, what variables should be logical/binary, factors, or numerical? And should the factor variables be ordered?*

`pclass`, `sex` and `embarked` should be factors.
```{r}
titanic <- titanic |>
  mutate(
    survived = survived == 1,
    pclass = factor(pclass, levels = c("1st", "2nd", "3rd"), ordered = TRUE),
    sex = factor(sex),
    embarked = factor(embarked)
  )
```

### Question 3
*Construct a new logical/binary variable `accompanied` to record if a passenger was accompanied by any relatives.*
```{r}
titanic <- titanic |>
  mutate(accompanied = (sibsp + parch) > 0)

titanic |> count(accompanied)
```

### Question 4
*Create plot(s) visualising the relative proportions of survival by different categories (similar to the above plot). Describe the relationships, do the relationships make sense?*

```{r}

```


